---
title: 'Figure 4'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='/local1/USERS/jfleck/data/PUBLIC_ORGANOIDS')
```

This notebook reproduces the main analyses from figure 4 of the manuscript. First we import the necessary packages.

```{r message=FALSE, warning=FALSE, results='hide'}
library(voxhunt)
library(tidyverse)
library(Seurat)
```

Now we load the data. The loaded seurat object contains the neuronal popultions of the datasets shown in the manuscript. We further subset the ones shown in figure 2.

```{r}
load_aba_data('voxhunt_data/')
neurons <- read_rds('combined_neurons_srt.rds')
neurons <- subset(neurons, orig.ident%in%c('cerebral', 'hCS', 'hSS', 'tanaka_thalamus'))
neurons <- subset(neurons, 
    cluster%in%c('mesen_ex_cerebral', 'mesen_in_cerebral', 'ctx_ex_cerebral', 
        'ge_in_cerebral', 'dien_ex_cerebral', 'ge_hss', 'ctx_hcs', 'dien_tho')
)
print(unique(neurons$cluster))
```

We can see that `cluster` already captures the different neuronal types we are interested in. Now we select some structure markers.

```{r}
struct_markers <- structure_markers('E13', 'custom_3')
genes_use <- struct_markers %>% 
    group_by(group) %>% 
    top_n(10, auc) %>% 
    pull(gene) %>% unique()
print(head(genes_use))
```

Now we run VoxHunt using these genes

```{r, fig.height=3, fig.width=5}
neuron_voxmap <- voxel_map(
    neurons, 
    group_name='cluster', 
    genes_use=genes_use
)
plot_map(neuron_voxmap) & no_legend()
```

As shown in the figure, we can also plot coronal slices. We can first pick the slices from the annotated map.

```{r, fig.width=8, fig.height=3}
voxhunt::plot_annotation('E13', show_coordinates = T, show_legend = T)
```

Now we plot slices 6, 11, 23 and 28

```{r, fig.height=7, fig.width=5}
voxhunt::plot_map(neuron_voxmap, view='slice', slices=c(6, 11, 23, 28)) & 
    no_legend()
```


```{r, include=F, results='hide'}
## Struct ape paper
struct_names <- c(
    'pallium',
    'subpallium',
    'preoptic telencephalon',
    'hypothalamus',
    'diencephalon',
    'midbrain',
    'hindbrain',
    'NA'
)

struct_colors <- c(
    '#ad1457',
    '#7b1fa2',
    '#5e35b1',
    '#ba68c8',
    '#303f9f',
    '#0097a7',
    '#43a047',
    'gray'
)
names(struct_colors) <- struct_names
```

Now we can also assign each cell to the highest correlating structure, similar as shown in figure 2.

```{r, fig.height=4, fig.width=6}
cell_assign <- assign_cells(neuron_voxmap)
cell_meta <- as_tibble(neurons@meta.data, rownames='cell') %>% 
    dplyr::select(-stage) %>% 
    dplyr::inner_join(cell_assign) %>% 
    # dplyr::filter(custom_2!='medullary hindbrain') %>% 
    dplyr::mutate(struct_name=case_when(
        str_detect(custom_2, 'hindbrain') ~ 'hindbrain',
        str_detect(custom_4, 'septum|subpall|striatum|amygda|telencephalic') ~ 'subpallium',
        str_detect(custom_2, 'telen') ~ 'pallium',
        TRUE ~ custom_2
    )) 

ggplot(cell_meta, aes(UMAP1, UMAP2, color=struct_name)) +
    geom_point(size=0.2) +
    facet_wrap(~dataset) +
    scale_color_manual(values=struct_colors) +
    theme_void()
```


```{r, fig.height=2, fig.width=4}
ggplot(cell_meta, aes(cluster, fill=struct_name)) +
    geom_bar(position='fill') +
    coord_flip() +
    scale_fill_manual(values=struct_colors)
```







