---
title: 'Figure 3'
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='/local1/USERS/jfleck/data/PUBLIC_ORGANOIDS')
```

This notebook reproduces the main analyses from figure 3 of the manuscript. First we import the necessary packages.

```{r message=FALSE, warning=FALSE, results='hide'}
library(voxhunt)
library(tidyverse)
library(Seurat)
```

Now we load the data. The loaded seurat object contains the neuronal popultions of the datasets shown in the manuscript. We further subset the ones shown in figure 2.

```{r}
load_aba_data('voxhunt_data/')
neurons <- read_rds('combined_neurons_srt.rds')
neurons <- subset(neurons, orig.ident%in%c('cerebral', 'hCS', 'hSS', 'tanaka_thalamus'))
neurons <- subset(neurons, 
    cluster%in%c('mesen_ex_cerebral', 'mesen_in_cerebral', 'ctx_ex_cerebral', 
        'ge_in_cerebral', 'dien_ex_cerebral', 'ge_hss', 'ctx_hcs', 'dien_tho')
)
print(unique(neurons$cluster))
```

We can see that `cluster` already captures the different neuronal types we are interested in. Now we select some structure markers.

```{r}
struct_markers <- structure_markers('E13', 'custom_2')
genes_use <- struct_markers %>% 
    group_by(group) %>% 
    top_n(15, auc) %>% 
    pull(gene) %>% unique()
print(head(genes_use))
```


```{r, include=F, results='hide'}
## Struct ape paper
struct_names <- c(
    'pallium',
    'subpallium',
    'preoptic telencephalon',
    'hypothalamus',
    'diencephalon',
    'midbrain',
    'hindbrain',
    'NA'
)

struct_colors <- c(
    '#ad1457',
    '#7b1fa2',
    '#5e35b1',
    '#ba68c8',
    '#303f9f',
    '#0097a7',
    '#43a047',
    'gray'
)
names(struct_colors) <- struct_names
```


Now we run VoxHunt with the selected genes and plot the similarities to brain structures.

```{r message=FALSE, warning=FALSE, fig.height=4, fig.width=6}
neuron_voxmap <- voxel_map(
    neurons, 
    group_name='cluster', 
    genes_use=genes_use, 
    pseudobulk_groups=T
)

strct_cor <- summarize_groups(neuron_voxmap) %>% 
    dplyr::filter(custom_2!='medullary hindbrain') %>% 
    dplyr::mutate(struct_name=case_when(
        str_detect(custom_2, 'hindbrain') ~ 'hindbrain',
        str_detect(custom_4, 'septum|subpall|striatum|amygda|telencephalic') ~ 'subpallium',
        str_detect(custom_2, 'telen') ~ 'pallium',
        TRUE ~ custom_2
    )) %>% 
    dplyr::mutate(struct_name=factor(struct_name, levels=struct_names)) %>% 
    dplyr::arrange(struct_name) %>% 
    dplyr::mutate(custom_4=factor(custom_4, levels=unique(.$custom_4))) %>% 
    dplyr::group_by(group, custom_4, struct_name) %>% 
    dplyr::summarise(corr=mean(corr)) %>% 
    dplyr::filter(!str_detect(custom_4, 'telencephalo')) 

ggplot(strct_cor, aes(custom_4, corr, fill=struct_name)) +
    geom_bar(stat='identity') +
    facet_wrap(group~., scales = 'free', ncol=2) +
    scale_fill_manual(values=struct_colors) +
    theme_bw() +
    theme(axis.text.x=element_blank(), axis.ticks.x=element_blank())
```

If we further run VoxHunt on a single cell level, we can assign each cell to the maximum correlating structure using `assign_cells()` we can then color the UMAP projection based on this assignment.





