---
title: "Figure 7"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir='/links/groups/treutlein/USERS/jfleck/data/PUBLIC_ORGANOIDS/')
```


This notebook reproduces the main analyses from figure 7 of the manuscript. First we import the necessary packages.

```{r, message=FALSE, warning=FALSE, results='hide'}
library(Seurat)
library(tidyverse)
library(patchwork)
library(voxhunt)
```

Now we set up some colors and load the data. The loaded seurat object contains the bulk RNA-seq data from the patterning screen. Each sample represents one organoid. We also do some feature selection with VoxHunt.

```{r}
data("voxel_meta")
load_aba_data('voxhunt_data/')
patscreen_data <- read_rds('patscreen_srt.rds')

ps_names <- c(
    'Ctrl_no_dose', 
    'SHH_high', 'SHH_low', 
    'CHIR_high', 'CHIR_low', 
    'RSPO2_high', 'RSPO2_low', 
    'RSPO3_high', 'RSPO3_low',
    'SB_DM_high', 'SB_DM_low'
)
ps_colors <- c(
    'gray', 
    '#f57f17', '#ffc107', 
    '#186a3b', '#28b463', 
    '#c2185b', '#f48fb1',
    '#512da8', '#9575cd',
    '#2874a6', '#5dade2'
)
names(ps_colors) <- ps_names

struct_markers <- structure_markers('E13', 'custom_2')
genes_use <- struct_markers %>% 
    group_by(group) %>% 
    top_n(10, auc) %>% 
    pull(gene) %>% unique()

print(patscreen_data)

```

Now we perform PCA to get an impression how the samples are distributed.


```{r, fig.height=4, fig.width=6}
patscreen_data <- patscreen_data %>% FindVariableFeatures(nfeatures=2000) %>% ScaleData() %>% RunPCA(npcs=5)
patscreen_data@reductions$pca@stdev
DimPlot(patscreen_data, group.by='mor_dose', pt.size=4) +
    scale_color_manual(values=ps_colors) +
    no_legend() 
```

Now we map the data with VoxHunt.

```{r, fig.height=6, fig.width=8}
ps_map <- voxel_map(patscreen_data, stage='E11', genes_use = genes_use, group_name = 'morphogen')
plot_map(ps_map)
```

Further, we'll have a closer look at the correlation patterns of RSPO2/3.

```{r, fig.height=3, fig.width=8}
rspo_data <- subset(patscreen_data, morphogen%in%c('RSPO2', 'RSPO3'))
ps_map <- voxel_map(rspo_data, stage='E13', genes_use = genes_use, group_name = 'morphogen')
plot_map(ps_map, slices = 6)
```

Lastly, we can look at some markers on the same slice to see that RSPO2 & 3 are expressed adjacent to the highlighted position.

```{r, fig.height=4, fig.width=6}
plot_expression('E13', slices = 6, genes = c('RSPO2', 'RSPO3', 'TTR', 'SPINT2'))
```

